### 2016-07-10 version0.0/intro0.0.tex
finishing intro0.0

\section{Prelude}
\label{sec: overview}
%what do we have inside galaxies:
% stuff we don't care about because they don't have a big contribution on the SED 1) planets; 2) black holes 3) dark matter
Inside every galaxy, there are millions of stars, planets, black holes, and high amount of dust and gas which all affect each others' formation, evolution and extinction.
There are also dark matter inside every galaxy, which to the best our knowledge only has gravitational effect on the galaxies.%add ref.s
All the information we can obtain from galaxies other than our owns are implanted in their light. % now we can detect gravitational wave which is not part of the SEDs, right?
Light emitted from young and evolved stars, and dust and gas in the space between stars have the most contribution on spectral energy distributions (SEDs) of galaxies.
We cannot detect any contribution from planets and black holes on SEDs due to their low intensity or the limitation of our detectors.
Therefore, with properly modeling observed features in SEDs we can extract information about evolution of stars, dust and gas inside galaxies.

@@

Formation of stars is a complicated process and directly affect on galaxies evolution. 
Various phenomena affect the collapse of molecular clouds and star formation which include the environment of the star-forming region, chemical compositions, the initial mass distribution of gas, gas accretion and cooling, H$_2$ formation, etc. 
For more than 50 years, the numerical simulations, the empirical and theoretical laws, which describe the star formation rate have been investigated; nevertheless, their results have never been precise. 
There are empirical scaling laws to relate the star formation to other properties of galaxies. 
The most well-known of these scaling laws is the relation between the surface density of the star formation rate (SFR) and the gas surface density, which is known as the Kennicutt-Schmidt (KS) law~\citep{Schmidt59, Kennicutt98a}. 
Similar to KS law, another scaling relation was derived which is a linear scaling between the total SFR and the mass of molecular gas (refer to reviews~\cite{Elmgrenn11, Kennicutt12}).

@@

 Studying formation of stars helps us understand the formation and evolution of the universe. 
 Since forming stars contains many complex physical and chemical processes, theoretical studies cannot yet give a complete picture of star formation.
 Therefore empirical studies are used to find laws that indicate the relations between physical properties of galaxies and star formation.
 Between different scaling relations for star formation, the the Kennicutt-Schmidt (KS) law \citep{Schmidt59, Kennicutt98a} is the most tested empirical star formation law. 
 Since 1959, many studies have tested KS law on regions inside the Milky Way, nearby galaxies and high redshift galaxies.
 Not only observational studies, but also  simulations both apply and investigate KS law. 
 Despite all these studies, there are still many unsolved questions about the KS law; Is it really a good correlation? What is the slope, and is it unique?, Does SFR surface density only scale with gas density? does the scatter affect by a second parameter(s)? What is the limit for applying the KS law? Is the SFR surface density correlation with total gas density stranger than with the H$_2$ surface density alone?

Many groups are studying the KS law from different approaches to answer the questions listed in above. The extended Schmidt law is a newer version of the KS law, which by considering the effect of old stellar population on the SFR provides stronger correlation for the SFR than the other scaling laws. Although in original paper this law is tested on large sample of various types of galaxies, but further studies are necessary, due to the fact that this law is tested on the total SFR and not on different regions on galaxies to determine effect of environment on the SFR. In this project, the extended version of the Schmidt law will be tested on the Andromeda galaxy using both pixel by pixel relation and radial average along the disk. To due so star formation rate will be measured using equations~\ref{equ: halphaplus24} and ~\ref{sfr_tot_IR}. In order to find the gas surface density, direct method to map the ISM is chosen. Stellar surface density will be calculated by calibrations from both \cite{Eskew12} and \cite{Zhu10}.

The Andromeda galaxy is chosen for this project due to its similarity in the structure to the MW. Also it is the nearest spiral galaxy and provides a clear image of a galaxy like our own with the advantage that we get a "global" picture, whereas, studies of the MW are limited by our location within the Galaxy.  The Andromeda is closest spiral galaxy to the MW, therefore, we can have higher resolution images of it than we have of any other spiral galaxy. Having high resolution images provides us data from different regions within the galaxy with different physical situations (e.g. metallicity, surface brightness and etc.). All these parameters make the Andromeda a suitable test bed for studying the scaling law. The KS law has been tested on the Andromeda galaxy many times but the results have been controversial. The range of slope of KS law calculated for the Andromeda galaxy is between 0.5 - 2 \citep[e.g.,][]{Tabatabaei10, Ford13}. Testing extended Schmidt law on the Andromeda galaxy would help us recognise parameters that affect the star formation in the Andromeda galaxy and leads us to model and simulate star formation more accurately. Also it helps us answer questions about in what types of the regions (e.g. in case of different surface brightness, metallicity, distance from the disk, etc.) effect of the stellar mass is more important.

This thesis focuses on effect of quantities of galaxies on SFR in nearby and high redshift galaxies.

>>>>

\section{Prelude}
\label{sec: overview}
%what do we have inside galaxies:
% stuff we don't care about because they don't have a big contribution on the SED 1) planets; 2) black holes 3) dark matter
Inside every galaxy, there are millions of stars, planets, black holes, and high amount of dust and gas which all affect each others' formation, evolution and extinction.
There are also dark matter inside every galaxy, which to the best our knowledge only has gravitational effect on the galaxies.
All the information we can obtain from galaxies other than our owns are implanted in their light. % now we can detect gravitational wave which is not part of the SEDs, right?
Light emitted from young and evolved stars, and dust and gas in the space between stars have the most contribution on spectral energy distributions (SEDs) of galaxies.
We cannot detect any contribution from planets and black holes on SEDs due to their low intensity or the limitation of our detectors.
Therefore, with properly modeling observed features in SEDs we can extract information about evolution of stars, dust and gas inside galaxies.

@@

@@

Complete picture of galaxies are formed and evolved are tied to unfolding the mystery of formation of stars.
 Various phenomena affect the collapse of molecular clouds and star formation which include environment of the star-forming region, chemical compositions of the region, initial mass distribution of gas, mass of sounding stars and other physical properties of galaxies
 For decades, Observational, theoretical and numerical studies have investigated star formation; nevertheless, their results have never been precise. 
 Empirical studies are used alongside of theoretical ones to find laws that indicate the relations between physical properties of galaxies and star formation.

The most well-known scaling law is the relation between the surface density of FR and the gas surface density, which is known as the Kennicutt-Schmidt (KS) law~\citep{Schmidt59, Kennicutt98a}. 
The K-S law and similar empirical laws were tested on local and high-redshift galaxies with various morphological types, on local and global scales and different ISM types~\citep[e.g][]{Kennicutt08,Bigiel08,Genzel10,Gnedin10,Shi11}.
Despite all these studies, many questions about star formation laws still remain unanswered; the limitations of star formation laws are yet undetermined and there is no definitive answer on whether there is any universal star formation laws.
Their dependence on physical properties of galaxies is nontrivial and sometimes differs by type of star forming regions.
Although relation between stars and gas are well established, the detail of this relation is not clear.
Not only different types of gas change SFR in different ways, but also variations in metallicity for each type alter it.
There is no fully sampled and universal IMF.
In fact there are many debates on weather IMF is universal or not.
Influence of evolved stars and distance of star forming regions from centre of galaxies on SFR is mentioned in studies but not fully investigated.
Besides studying the impact of mentioned properties of galaxies on SFR, searching for other properties of galaxies/star forming regions is continuing.

Advances in observing facilities, image qualities, and methodology expands our knowledge of formation of stars.
Besides these advances, in the past few decades,using advanced statistical methods grew in astronomical studies.
Applying modern statistical methods on star formation studies can shed a light on unsolved problems in this field. 
Studies show that using different statistical methods changes our perspective on star formation laws. 
For example,~\cite{Shetty13} used Bayesian methods to show the K-S law is not universal. 
On the other hand, since astronomy is a data driven science, data mining methods are quiet useful to get new insight on properties of star forming regions.

In this thesis, we present three projects that study effects of quantities of galaxies on SFR.
We use UV to radio observations of both nearby and high-redshift galaxies, measure their physical properties and using modern statistical methods study their impact on SFR.
Chapter~\ref{ch: paper1} focuses on investigating star formation laws in the Andromeda galaxies (M31) in both local and global scales using hierarchical Bayesian analysis.
Additional to the K-S law, we study the extended Schmidt law~\citep{Shi11} and and the metallicity/star formation correlation~\citep{Krumholz09}.
We also show the effect of using different SFR and ISM gas tracers on star formation laws.
In Chapter~\ref{ch: paper3}, we explore nearby galaxies using an unsupervised data mining method (self organizing map, SOM).
We cluster observational data and derived quantities of spatially resolved regions in M31 and examine the physical properties of each cluster.
We find hidden structures in high dimensional data space during this process.
In Chapter~\ref{ch: paper2}, we move to higher redshift and apply SOM algorithm on SED of high-redshift galaxies.
We train networks using a set of galaxy SED templates covering the wavelength range from UV to IR and use them to classify the SEDs of a sample of 142 galaxies with $0.5 < z < 1$. 
Then, we study the grouped properties of classified galaxies.
For each chapter, we also describe the methodology and data used in that chapter.
We summarize our results and discuss potential future work and applications of the results of this thesis in Chapter~\ref{ch: summary}
